---
title: "Exercise 05"
author: "Jennie"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Challenge 1

## Step 1

-   Using the {tidyverse} `read_csv()` function, load the “IMDB-movies.csv” dataset from [this URL](https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv) as a “tibble” named **d**

```{r Step1.1}
library(tidyverse)
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv") 
as_tibble(d)
```

## Step 2

-   Use a one-line statement to filter the dataset to include just movies from 1920 to 1979 and movies that are between 1 and 3 hours long (**runtimeMinutes** \>= 60 and **runtimeMinutes** \<= 180), and add a new column that codes the **startYear** into a new variable, **decade** (“20s”, “30s”, …“70s”). If you do this correctly, there should be 5651 movies remaining in the dataset.

```{r Step1.2}
library(dplyr)
d <- d |> filter(runtimeMinutes>=60 & runtimeMinutes<=180) |> filter(startYear>=1920 & startYear<=1979) |> mutate(Decade = if_else(startYear >= 2000, paste0(startYear  %/% 10 * 10, "s"), paste0((startYear - 1900) %/% 10 * 10, "s")))
```

## Step 3

-   Use {ggplot2} (which is part of {tidyverse}) to plot histograms of the distribution of **runtimeMinutes** for each decade.

```{r Step 1.3}
library(ggplot2)
plots <- ggplot(d, aes(x=runtimeMinutes)) + 
  geom_histogram()
plots + facet_wrap(vars(Decade), nrow = 2)
```

## Step 4

-   Use a one-line statement to calculate the population mean and population standard deviation in **runtimeMinutes** for each decade and save the results in a new dataframe called **results**.

```{r Step1.4}
results <- d |> group_by(Decade) |> summarize(mean = mean(runtimeMinutes), sd = sd(runtimeMinutes))
```

## Step 5

-   Draw a single sample of 100 movies, without replacement, from each decade and calculate the single sample mean and single sample standard deviation in **runtimeMinutes** for each decades. Recall that your single sample mean for each decade is an *estimate* of the population mean for each decade.

```{r Step1.5}
d_sample <- d |> group_by(Decade) |> slice_sample(n = 100, replace = FALSE) 
d_sample_summary <- d_sample |> summarize(mean = mean(runtimeMinutes), sd = sd(runtimeMinutes))
```

## Step 6

-   Calculate for each decade the standard error around your estimate of the population mean **runtimeMinutes** based on the standard deviation and sample size (n=100 movies) of your single sample.

```{r Step1.6}
d_sample_summary$se <- (d_sample_summary$sd/sqrt(100))
```

## Step 7

-   Compare these estimates to the actual population mean **runtimeMinutes** for each decade and to the calculated SE in the population mean for samples of size 100 based on the population standard deviation for each decade.

```{r Step1.7}
# difference in population and sample means
d_sample_summary$mean_diff <- (results$mean - d_sample_summary$mean)

# SE from pop is population sd divided by sample size
d_sample_summary$se_pop <- (results$sd/sqrt(100))

# difference in SE 
d_sample_summary$se_diff <- (d_sample_summary$se_pop - d_sample_summary$se)

```

## Step 8

-   Generate a *sampling distribution* of mean **runtimeMinutes** for each decade by [a] drawing 1000 random samples of 100 movies from each decade, without replacement, and, for each sample, [b] calculating the mean **runtimeMinutes** and the standard deviation in **runtimeMinutes** for each decade. Use either a standard `for( ){ }` loop, the `do(reps) *` formulation from {mosaic}, the `rerun()` function from {purrr}, or the `rep_sample_n()` workflow from {infer} to generate your these sampling distributions (see [**Module 16**](https://difiore.github.io/ada-2024/16-module.html)).

```{r Step1.8}
library(infer)
reps <- 1000
samp_dist_decade <- d |> rep_sample_n(size = 100, reps = reps, replace = FALSE) |> group_by(Decade) |> summarize(mean = mean(runtimeMinutes), sd = sd(runtimeMinutes))

```

## Step 9

-   Then, calculate the **mean** and the **standard deviation** of the sampling distribution of sample means for each decade (the former should be a very good estimate of the population mean, while the latter is another estimate of the standard error in our estimate of the population mean for a particular sample size) and plot a histogram of the sampling distribution for each decade. What shape does it have?

```{r Step 1.9}

```

## Step 10

-   Finally, compare the standard error in **runtimeMinutes** for samples of size 100 from each decade [1] as estimated from your **first** sample of 100 movies, [2] as calculated from the known *population* standard deviations for each decade, and [3] as estimated from the sampling distribution of sample means for each decade.

```{r Step 1.10}

```

# Challenge 2

## Step 1
