---
title: "Exercise 05"
author: "Jennie"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Challenge 1

## Step 1

-   Using the {tidyverse} `read_csv()` function, load the “IMDB-movies.csv” dataset from [this URL](https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv) as a “tibble” named **d**

```{r Step1.1}
library(tidyverse)
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv") 
as_tibble(d)
```

## Step 2

-   Use a one-line statement to filter the dataset to include just movies from 1920 to 1979 and movies that are between 1 and 3 hours long (**runtimeMinutes** \>= 60 and **runtimeMinutes** \<= 180), and add a new column that codes the **startYear** into a new variable, **decade** (“20s”, “30s”, …“70s”). If you do this correctly, there should be 5651 movies remaining in the dataset.

```{r Step1.2}
library(dplyr)
d <- d |> filter(d$runtimeMinutes>=60 & d$runtimeMinutes<=180) |> filter(d$startYear>=1920 & d$startYear<=1979) |> mutate(Decade = if_else(startYear >= 2000, paste0(startYear  %/% 10 * 10, "'s"), paste0((startYear - 1900) %/% 10 * 10, "'s")))
```

## Step 3

-   Use {ggplot2} (which is part of {tidyverse}) to plot histograms of the distribution of **runtimeMinutes** for each decade.

```{r Step 1.3}
library(ggplot2)
plots <- ggplot(d, aes(x=runtimeMinutes)) + 
  geom_histogram()
plots + facet_wrap(vars(Decade), nrow = 2)
```

## Step 4

-   Use a one-line statement to calculate the population mean and population standard deviation in **runtimeMinutes** for each decade and save the results in a new dataframe called **results**.

```{r Step1.4}
results <- d |> group_by(Decade) |> summarize(mean = mean(runtimeMinutes), sd = sd(runtimeMinutes))
```

## Step 5

-   Draw a single sample of 100 movies, without replacement, from each decade and calculate the single sample mean and single sample standard deviation in **runtimeMinutes** for each decades. Recall that your single sample mean for each decade is an *estimate* of the population mean for each decade.

```{r Step1.5}
d_sample <- d |> group_by(Decade) |> slice_sample(n = 100, replace = FALSE) |> summarize(mean = mean(runtimeMinutes), sd = sd(runtimeMinutes))
```

## Step 6

-   Calculate for each decade the standard error around your estimate of the population mean **runtimeMinutes** based on the standard deviation and sample size (n=100 movies) of your single sample.

```{r Step1.6}

```